Practical Optional Types for Clojure

### 2015-02-25 overview.tex
hmap syntax by example

\begin{exmp}
\begin{listing}
\inputminted[firstline=5,lastline=8]{clojure}{code/demo/src/demo/hmap.clj}
\inputminted[firstline=26,lastline=28]{clojure}{code/demo/src/demo/hmap.clj}
\end{listing}
\end{exmp}

@@

HMaps support various options to track the presence and absence of entries.
\clj{'{:a Num}} is a shorthand for the \clj{:mandatory} option---entries that
must be present:
$$
\clj{(HMap :mandatory {:a Num})}.
$$

\clj{:absent-keys} is a set of keywords that are known to not have entries.
The value \clj{{:a 1 :b nil}} is not a member of the type
% '#' not allowed in mintinline
\smallskip
\begin{minted}{clojure}
  (HMap :mandatory {:a Num} :absent-keys #{:b}).
\end{minted}

\clj{:optional} is a map that declares entries either absent or
present with a specific type. Concretely
$$
\clj{(HMap :optional {:a Num})}
$$
is the same as
\begin{minted}{clojure}
      (U '{:a Num} (HMap :absent-keys #{:a})).
\end{minted}

@@

\clj{:mandatory}, \clj{:optional} and \clj{:absent-keys} default to empty
and \clj{:complete?} defaults to \clj{false}.

@@

\begin{exmp}
\begin{listing}
\inputminted[firstline=5,lastline=9]{racket}{code/demo/src/demo/hmap_path.clj}
\end{listing}
\end{exmp}

@@

\begin{exmp}
\begin{listing}
\inputminted[firstline=15,lastline=16]{clojure}{code/demo/src/demo/parent3.clj}
\end{listing}
\end{exmp}

>>>>

\begin{Code}
\begin{exmp}
\inputminted[firstline=6,lastline=8]{clojure}{code/demo/src/demo/hmap.clj}
\inputminted[firstline=26,lastline=28]{clojure}{code/demo/src/demo/hmap.clj}
\end{exmp}
\end{Code}

@@

HMaps support various options to track the presence and absence of entries.
\clj{'{:a Num}} is a shorthand for the \clj{:mandatory} option---entries that
must be present.

\begin{Code}
\begin{exmp}
\inputminted[firstline=58,lastline=64]{clojure}{code/demo/src/demo/hmap.clj}
\end{exmp}
\end{Code}

Looking up \clj{:a} simply returns the associated type.
Keys other than \clj{:a} are unaccounted for in \clj{Mand}, so 
the best we can infer is \clj{Any}, the supertype of all types.

\clj{:absent-keys} is a set of keywords that are known to not have entries.
We take advantage of this to infer a better type for lookups on known absent keys.

\begin{Code}
\begin{exmp}
\inputminted[firstline=66,lastline=73]{clojure}{code/demo/src/demo/hmap.clj}
\end{exmp}
\end{Code}

\clj{:optional} is a map that declares entries either absent or
present with a specific type.

\begin{Code}
\begin{exmp}
\inputminted[firstline=5,lastline=12]{clojure}{code/demo/src/demo/hmap_path.clj}
\end{exmp}
\end{Code}

Concretely \clj{(HMap :optional {:a Num})}
is the same as
\begin{minted}{clojure}
      (U '{:a Num} (HMap :absent-keys #{:a})).
\end{minted}

@@

\begin{Code}
\begin{exmp}
\inputminted[firstline=17,lastline=23]{clojure}{code/demo/src/demo/hmap_path.clj}
\end{exmp}
\end{Code}

Typed Clojure knows that this \clj{merge} will preserve the entries
of the left hand map.

\begin{Code}
\begin{exmp}
\inputminted[firstline=25,lastline=25]{clojure}{code/demo/src/demo/hmap_path.clj}
\end{exmp}
\end{Code}

@@

\begin{Code}
\begin{exmp}
\inputminted[firstline=14,lastline=15]{clojure}{code/demo/src/demo/hmap_path.clj}
\end{exmp}
\end{Code}

@@

\begin{Code}
\begin{exmp}
\inputminted[firstline=15,lastline=16]{clojure}{code/demo/src/demo/parent3.clj}
\end{exmp}
\end{Code}
